Supplement to ``{\mytitle}''

### 2022-10-12 ms.Rnw
Worked on references

As previously stated, Model~2 is a deterministic compartmental model defined by a set of coupled differential equations.
The use of deterministic compartment models have a long history in the field of infectious disease epidemiology \citep{kermack1927,brauer2017,varghese21},
and can be justified by asymptotic considerations in a large-population limit \citep{dadlani2020,ndii17}.
Because the process model of Model~2 is deterministic, the parameter estimation problem for Model~2 reduces to a least squares calculation when combined with a Gaussian measurement model (see supplement material for details).
\cite{lee20} fit two versions of model~2 based on a presupposed change in cholera transmission from a epidemic phase to endemic phase that occurred in March, 2014.
The inclusion of a change-point in model states and parameters increases the flexibility of the model and hence the ability to fit the observed data.
The increase in model flexibility, however, results in hidden states that are inconsistent between model phases.
% A naive solution to this would be to estimate only model parameters in separate phases and not the hidden states, but this may also result in hidden states in one phase of the model that are inconsistent with the other phase of the outbreak, since only the data from one phase is used to fit model parameters.
The inclusion of a model break-point by \cite{lee20} is perhaps due to a challenging feature of fitting a deterministic model via least squares: discrepancies between model trajectories and observed case counts in highly infectious periods of a disease outbreak will result in greater penalty than the discrepancies between model trajectories and observed case counts in times of relatively low infectiousness, resulting in a bias towards accurately describing periods of high infectiousness.
This issue is particularly troublesome for modeling cholera dynamics in Haiti: the inability to accurately fit times of low infectiousness may result in poor model forecasts, as few cases of cholera were observed the last few years of the epidemic.

>>>>

As previously stated, Model~2 is a deterministic compartmental model defined by a set of coupled differential equations.
The use of deterministic compartment models have a long history in the field of infectious disease epidemiology \citep{kermack1927,brauer2017,giordano20},
and can be justified by asymptotic considerations in a large-population limit \citep{dadlani2020,ndii17}.
Because the process model of Model~2 is deterministic, the parameter estimation problem for Model~2 reduces to a least squares calculation when combined with a Gaussian measurement model (see supplement material for details).
\cite{lee20} fit two versions of model~2 based on a presupposed change in cholera transmission from a epidemic phase to endemic phase that occurred in March, 2014.
The inclusion of a change-point in model states and parameters increases the flexibility of the model and hence the ability to fit the observed data.
The increase in model flexibility, however, results in hidden states that are inconsistent between model phases.
% A naive solution to this would be to estimate only model parameters in separate phases and not the hidden states, but this may also result in hidden states in one phase of the model that are inconsistent with the other phase of the outbreak, since only the data from one phase is used to fit model parameters.
The inclusion of a model break-point by \cite{lee20} is perhaps due to a challenging feature of fitting a deterministic model via least squares: discrepancies between model trajectories and observed case counts in highly infectious periods of a disease outbreak will result in greater penalty than the discrepancies between model trajectories and observed case counts in times of relatively low infectiousness, resulting in a bias towards accurately describing periods of high infectiousness.
This issue is particularly troublesome for modeling cholera dynamics in Haiti: the inability to accurately fit times of low infectiousness may result in poor model forecasts, as few cases of cholera were observed the last few years of the epidemic.

### 2022-10-12 si/si.Rnw
Worked on references

In this model, the advantage afforded by vaccination is an increased probability that an infection is asymptomatic.
Therefore, under the assumptions of this model, all reported cases are assumed to be a fraction of individuals that transition from the exposed to the infected compartment, as noted in Eq.~\myeqref{model1:meas}:
\begin{equation}
  \label{model1:meas}
  Y_{n} \mid \Delta N_{EI}(n) = z(n) \sim \text{NB}\left(\reportRate z(n), \obsOverdispersion \right),
\end{equation}
where $Y_n$ is the reported cholera cases at time $n \in 1:N$ and $\Delta N_{EI}(n)$ is the sum total of individuals across vaccination compartment $\vaccCounter$ who moved from compartment $E_\vaccCounter$ to $I_\vaccCounter$ since observation $n-1 \in 0:N-1$, and $\text{NB}\left(\reportRate z, \obsOverdispersion \right)$ is the negative binomial distribution with mean $\reportRate z$ and variance $\reportRate z+ \psi^2\reportRate^2 z^2)$. \jwc{CHECK THIS}.

>>>>

In this model, the advantage afforded by vaccination is an increased probability that an infection is asymptomatic.
Therefore, under the assumptions of this model, all reported cases are assumed to be a fraction of individuals that transition from the exposed to the infected compartment, as noted in Eq.~\myeqref{model1:meas}:
\begin{equation}
  \label{model1:meas}
  Y_{n} \mid \Delta N_{EI}(n) = z(n) \sim \text{NB}\left(\reportRate z(n), \obsOverdispersion \right),
\end{equation}
where $Y_n$ is the reported cholera cases at time $n \in 1:N$ and $\Delta N_{EI}(n)$ is the sum total of individuals across vaccination compartment $\vaccCounter$ who moved from compartment $E_\vaccCounter$ to $I_\vaccCounter$ since observation $n-1 \in 0:N-1$, and $\text{NB}\left(\reportRate z, \obsOverdispersion \right)$ is the negative binomial distribution with mean $\reportRate z$ and variance $\reportRate z \big(1 + \frac{\reportRate z}{\obsOverdispersion}\big)$.
